### 2017-10-19 submitted-v1/cover-letter.tex
submission version 1

\documentclass{letter}
\usepackage{graphics}
\usepackage{fullpage}
\topmargin=-0.2in
%\headheight=0in
\pdfminorversion=4 % as instructed by JASA file upload
\setlength\oddsidemargin{-0.2in}
\textheight=8.5in
\textwidth=7in
\def\apl#1#2{
\begin{letter}{#1}
\signature{Mahbubul Majumder, PhD.\\ Assistant Professor \\ Department of Mathematics \\University of Nebraska at Omaha \\ Omaha, NE 68118 \\ Phone: (402) 554 2734 \\ {\tt mmajumder@unomaha.edu} \\ with Dr. Heike Hoffman, Iowa State university and Dianne Cook, Monash University, Australia}
\opening{{Dear #2,}}

Please find enclosed a manuscript  ``Human Factors Influencing Visual Statistical Inference'' which we are
submitting for consideration for publication in the
Applications and Case Studies section of the Journal of American Statistical
Association.

This paper is the complement of the two foundational papers describing research which bridges
graphics, exploratory data analysis/data mining, and statistical
inference. The first paper by Buja, et al (2009) published in RSS Phil. Trans. A,
sets up the conceptual foundation of visual inference. The second paper by Majumder, et al (2013) published in JASA Theory and applications section, clarifies and defines the approach and validates it with a user study. It was one of the most read JASA articles and selected to be a feature article. This paper studies the human factors that may influence the visual inference conducting a series of 10 experiments with human subjects. We believe this paper is of interest to the JASA readership.

Thank you for your consideration of our work.

\closing{Sincerely yours,
}
\end{letter}}

\begin{document}

{{\pdfimage width 7in {letterhead.png}}}
\apl{Professor Montserrat Fuentes\\Applications and Case Studies Editor\\Journal of American Statistical Association}{Professor Fuentes}

\end{document}